### 2016-05-03 ms_oviemission.tex
add context about simulation in introduction

\documentclass[useAMS,usenatbib]{mn2e}
\input psfig.sty
\usepackage{array}
%\usepackage{graphicx}
%\usepackage{epstopdf}
\usepackage{amsmath}

\voffset=-0.8in

\def \gas {\textsc{gasoline}}
\def \gastwo {\textsc{ESF-gasoline2}}
\def \changa {\textsc{changa}}
\def \mean#1{\left< #1 \right>}

%%MISC
\def \aj {AJ}
\def \apj {ApJ}
\def \apjl {ApJL}
\def \mnras {MNRAS}
\def \apjs {ApJS}
\def \aap {A\&A}
\def \nat {Nature}
\def \pasp {PASP}
\def \na {NewA}

\def \etal {et~al.~}
\def \eg{e.g.}
\def \Section{\S}
\def \spose#1{\hbox  to 0pt{#1\hss}}  
\def \lta{\mathrel{\spose{\lower 3pt\hbox{$\sim$}}\raise  2.0pt\hbox{$<$}}}
\def \gta{\mathrel{\spose{\lower  3pt\hbox{$\sim$}}\raise 2.0pt\hbox{$>$}}}
\def \ion#1#2{#1{\footnotesize{#2}}\relax}
\def \ha       {H$\alpha$}
\def \hi       {\ion{H}{I}}
\def \hii      {\ion{H}{II}}
\def \oii      {[\ion{O}{II}]}
\def \ovi      {\ion{O}{VI}}

%%UNITS
\def \kmsmpc {\>{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
\def \kms {\ifmmode  \,\rm km\,s^{-1} \else $\,\rm km\,s^{-1}  $ \fi }
\def \kpc {\ifmmode  {\rm kpc}  \else ${\rm  kpc}$ \fi  }  
\def \hkpc {\ifmmode  {h^{-1}\rm kpc}  \else ${h^{-1}\rm kpc}$ \fi  }  
\def \hMpc {\ifmmode  {h^{-1}\rm Mpc}  \else ${h^{-1}\rm Mpc}$ \fi  }  
\def \Msun {\ifmmode \rm M_{\odot} \else $\rm M_{\odot}$ \fi}
\def \hMsun {\ifmmode h^{-1}\,\rm M_{\odot} \else $h^{-1}\,\rm M_{\odot}$ \fi}
\def \hhMsun {\ifmmode h^{-2}\,\rm M_{\odot}\else $h^{-2}\,\rm M_{\odot}$ \fi}
\def \Lsun {\ifmmode L_{\odot} \else $L_{\odot}$ \fi} 
\def \hhLsun {\ifmmode h^{-2}\,\rm L_{\odot} \else $h^{-2}\,\rm L_{\odot}$ \fi}

%%COSMOLOGY

\def\LCDM{$\Lambda$CDM }
\def \LCDM {\ifmmode \Lambda{\rm CDM} \else $\Lambda{\rm CDM}$ \fi}
\def \sig8 {\ifmmode \sigma_8 \else $\sigma_8$ \fi} 
\def \Omegam {\ifmmode \Omega_{\rm m} \else $\Omega_{\rm m}$ \fi} 
\def \Omegab {\ifmmode \Omega_{\rm b} \else $\Omega_{\rm b}$ \fi} 
\def \Omegar {\ifmmode \Omega_{\rm r} \else $\Omega_{\rm r}$ \fi} 
\def \fbar {\ifmmode f_{\rm bar} \else $f_{\rm bar}$ \fi} 
\def \OmegaL {\ifmmode \Omega_{\rm \Lambda} \else $\Omega_{\rm \Lambda}$\fi} 
\def \Deltavir {\ifmmode \Delta_{\rm vir} \else $\Delta_{\rm vir}$ \fi}
\def \rhocrit {\ifmmode \rho_{\rm crit} \else $\rho_{\rm crit}$ \fi}

%DARK MATTER
\def \rs {\ifmmode r_{\rm s} \else $r_{\rm s}$ \fi} 
\def \rh {\ifmmode r_{\rm h} \else $r_{\rm h}$ \fi} 
\def \Rvir {\ifmmode R_{\rm vir} \else $R_{\rm vir}$ \fi}
\def \Vvir {\ifmmode V_{\rm  vir} \else  $V_{\rm vir}$  \fi} 
\def \Vmax {\ifmmode V_{\rm  max} \else  $V_{\rm max}$  \fi} 
\def \Mvir {\ifmmode M_{\rm  vir} \else $M_{\rm  vir}$ \fi}  
\def \Mhalo {\ifmmode M_{\rm halo} \else $M_{\rm  halo}$ \fi}  
\def \Nvir {\ifmmode N_{\rm  vir} \else $N_{\rm  vir}$ \fi}  
\def \Jvir {\ifmmode J_{\rm vir} \else $J_{\rm vir}$ \fi} 
\def \Evir {\ifmmode E_{\rm vir} \else $E_{\rm vir}$ \fi} 
\def \lam {\ifmmode \lambda  \else $\lambda$ \fi} 
\def \lamp {\ifmmode \lambda^{\prime} \else $\lambda^{\prime}$  \fi} 
\def \lampc {\ifmmode \lambda^{\prime}_{\rm c} \else
  $\lambda^{\prime}_{\rm c}$  \fi}

\def \xoff {\ifmmode x_{\rm off} \else $x_{\rm off}$ \fi}
\def \rhorms {\ifmmode \rho_{\rm rms} \else $\rho_{\rm rms}$ \fi}
\def \qbar {\ifmmode \bar{q} \else $\bar{q}$ \fi}

%%BARYONS
\def \Mb {\ifmmode M_{\rm b} \else $M_{\rm b}$ \fi} 
\def \eSF {\ifmmode \epsilon_{\rm SF} \else $\epsilon_{\rm SF}$ \fi} 
\def \Md {\ifmmode M_{\rm d} \else $M_{\rm d}$ \fi} 
\def \Mg {\ifmmode M_{\rm g} \else $M_{\rm g}$ \fi} 
\def \Rb {\ifmmode R_{\rm b} \else $R_{\rm b}$ \fi} 
\def \Rd {\ifmmode R_{\rm d} \else $R_{\rm d}$ \fi} 
\def \Rg {\ifmmode R_{\rm g} \else $R_{\rm g}$ \fi} 
\def \mgal {\ifmmode m_{\rm gal} \else $m_{\rm gal}$ \fi} 
\def \rj {\ifmmode {\cal R}_j \else ${\cal R}_j$ \fi} 
\def \lamgal {\ifmmode \lambda_{\rm gal} \else $\lambda_{\rm gal}$ \fi} 
\def \Vcirc {\ifmmode V_{\rm circ} \else $V_{\rm circ}$ \fi} 
\def \Vrot {\ifmmode V_{\rm rot} \else $V_{\rm rot}$ \fi} 
\def \Vflat {\ifmmode V_{\rm flat} \else $V_{\rm flat}$ \fi} 
\def \Mstar {\ifmmode M_{\rm star} \else $M_{\rm star}$ \fi} 
\def \Mgas {\ifmmode M_{\rm gas} \else $M_{\rm gas}$ \fi} 
\def \Mbar {\ifmmode M_{\rm bar} \else $M_{\rm bar}$ \fi}

%%MASS-TO-LIGHT RATIOS
\def \DeltaIMF {\ifmmode \Delta_{\rm IMF} \else $\Delta_{\rm IMF}$ \fi}

\def \VV {\ifmmode V_{\rm 2.2}/V_{200} \else $V_{2.2}/V_{200}$ \fi} 
\def \dvr {\ifmmode \partial_{\rm VR} \else $\partial_{\rm VR}$ \fi}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[\hi{} and \ovi{} Emssivity] {\hi{} and \ovi{} Emission from the Circum-Galactic Medium
 with a large sample of cosmological hydrodynamical simulations}

\author[Zhang et al.]{Shuinai Zhang$^{1}$\thanks{snzhang@pmo.ac.cn}, Liang Wang$^{1,3}$,Taotao Fang$^{2}$, Li Ji$^{1}$,Thales Gutcke$^3$, 
\newauthor{Andrea V. Macci\`o$^{3,4}$, Xi Kang$^{1}$}\\
$^1$Purple Mountain Observatory, 2 West Beijing Road, Nanjing 210008, China\\
$^2$Xiamen Univeristy, Xiamen, China \\
$^3$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany\\
$^4$New York University Abu Dhabi, PO Box 129188, Abu Dhabi, UAE}
\begin{document}

\date{to be submitted to }

\pagerange{\pageref{firstpage}--\pageref{lastpage}}\pubyear{2016}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  We use the NIHAO galaxy formation simulations to study ultra-violet (UV)
  emission from circum-galactic medium (CGM) in galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses. We analyze the spatially-extended structures of emission lines
  from \ovi{} and \hi{}.
\end{abstract}

\begin{keywords}
  galaxies: evolution -- galaxies: formation -- galaxies: dwarf -- galaxies: spiral -- 
  methods: numerical -- cosmology: theory
\end{keywords}

\setcounter{footnote}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 1: INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

A complete galaxy formation model does not only require the understanding
of galactic components, such as stars and interstellar medium (ISM),
but also the gaseous medium surrounded galaxies and extended to virial
radius and beyond, the circumgalactic medium (CGM) namely.
The galactic components dominate the energy output of the system,
and always be focused by baryon counting \citep{Bell03}.
Compared to the cosmological baryonic fraction $\Omega_b/\Omega_m$,
stars and ISM come up significantly short on baryons. 
If we include the hot X-ray halo gas \citep{Baldry08,Yang09,
McGaugh10,Anderson10,Gupta12,Fang13}, the fracion is still under
the prediction from the cosmic baryonic fraction.

Galaxies have possibility apparently to exhibit a diffuse baryonic
content within CGM ({\bf citation}) which is observed via UV absorption
lines and has much lower temperature and density \citep{Werk13}.
Such CGM cannot be traced directly by X-ray emission or any other
radiative emission process.
The installation of the Cosmic Origins Spectrograph (COS) greatly
increased the capability of UV spectroscopy on the Hubble space
Telescope (HST) platform (e.g. Savage et al. 2010), enabling the
use of fainter quasar targets illuminating foreground galactic
haloes.
COS-Halos \citep{Tumlinson11, Thom12, Werk12, Werk13} has been 
working to investigate 
the properties of this elusive multiphase medium.
One of COS-Halos results demonstrated CGM is a large reservior
of baryons extanding far beyond the stellar disk of the galaxy.
This significant reservoir of halo gas clearly plays a central 
role in the cycling of baryons into and our of a galaxy, 
regulating this growth.

A number of large volume simulations \citep{Ford13, Ford15, Suresh15, Oppenheimer16} and 
zoom-in cosmological simulations \citep{Stinson12,Hummels13,Shull14} have 
investigated the properties of CGM and the link between 
galaxies and their metal-enriched CGM.
\citet{Ford15} do not include Active Galactic Nucleus (AGN) and do not
reproduce the \ovi{} bimodality, while \citet{Suresh15b} argue that
AGN feedback can remove the CGM around passive galaxies then reduce
the \ovi{} column densities around  passive galaxies.
\citet{Gutcke16} compared the column density profiles of \ovi{} and \hi{}
in CGM of galaxies of the cosmologycal hydrodynamical simulation suite 
NIHAO \citep{Wang15} with observations, studied the covering fraction of 
dense \hi{}, looked at the shape of the CGM and its chemical composition.
The simulations have a good agreement with observations and show observers
and modellers can compare with them in future studies.

Absorption line measurements have several limitations. 
Firstly, suitable background sources are rare so that there is 
typically only one or no background sight line per foreground
galaxy({\bf citation}). Because the area covered scales as the 
square of the impact parameter, the resulting measurements 
preferentially probe large impact parameters from foreground 
galaxies.
Finally, because of the 1D nature of absorption line constraints
it is often unclear what physical structure are probed.
Intergral field measurement of emission lines are complementary 
to absorption line measurements.
In particular, they allow us to directly obtain a 3D picture
of the distribution of gas in the CGM by combining the 2D map of
emission on the sky with line-of-sight velocity information.

\citet{Sravan15} studied ultra-violet (UV) metal line emission from the CGM
of high-redshift ($z=2-4$) galaxies using cosmological simulations from the 
Feedback in Realistic Environments (FIRE) project.
For each simulation, they predicted the emission from multiple metal lines
including \hi{} and \ovi{}, and found more massive haloes are on average 
more UV-luminous and the UV metal line emission is primarily from collisionally
ionized gas and strongly time variable. The high UV-luminous event correspond
high star formation and \ion{C}{III}, \ion{Si}{III}, \ion{C}{IV} and \ion{Si}{IV}
should be detectable by current and upcoming intergral field spectrographs such
as the Multi Unit Spectroscopic Explorer (MUSE{\bf citation}).

In this paper we investigate the CGM \ovi{} and \hi{} emissivity distribution
and shape in Numerical Investigation of a Hundred Astrophysical Objects,
NIHAO \citep{Wang15} project.
NIHAO project produced nearly 100 hydrodynamical, cosmological zoom-in galaxies
across a range in mass from

This paper is organized as follows: The cosmological hydrodynamical
simulations and the methodology for computing metal line emission
are briefly described in  \S\ref{sec:method}; In \S\ref{sec:results} 
we present the results including the \hi{} and \ovi{} emission map,
surface brightness and luminosity evolutions of all galaxies in NIHAO
sample; \S\ref{sec:sum} gives discussion and  summary of our
results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 2 METHODOLOGY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology} 
\label{sec:method}

\subsection{Simulations}
\label{sec:sims}

The simulations studied in this work are from the NIHAO project \citep{Wang15}.  
The halos to be re-simulated with baryons have been extracted from 
3 different pure N-body simulations with a box size of 60, 20 and 
15 $h^{-1}$ Mpc respectively\citet{Dutton14}. 
All halos across the whole mass range with typically a million dark
matter particles inside the virial radius of the target halo at 
redshift $z=0$.  
We adopted the  latest compilation of cosmological
parameters from the Planck  satellite \citep{Planck14}.

We use the SPH hydrodynamics code {\sc gasoline} \citep{Wadsley04},
with a revised treatment of  hydrodynamics as described in
\citet{Keller14}.  The code includes a subgrid model for turbulent
mixing of metal and energy \citep{Wadsley08}, heating and cooling
include photoelectric heating of dust grains, ultraviolet (UV) heating
and ionization and  cooling due to hydrogen, helium and metals
\citep{Shen10}.  The star formation and feedback modeling follows what
was used in the MaGICC simulations \citep{Stinson13}.  There are two
small changes in NIHAO simulations: The change in  number of neighbors
and the new combination of softening length and  particle mass means
the threshold for star formation increased from  9.3 to 10.3
cm$^{-3}$, the increase of pre-SN feedback efficiency $\epsilon_{\rm
  ESF}$, from 0.1 to 0.13.  The more detail on star formation and
feedback modeling can be found in \citet{Wang15}.

\subsection{Emissivity Calculation}
\label{sec:emis}

We first assign the hydrogen number densities and temperatures of all gas 
particles inside $2 \Rvir$ to $200\times 200 \times 200$ grids 
according SPH spline kernel \citep{Monaghan85}:
\begin{equation}
W\left(r,h\right) = \frac{8}{\pi h^3}
  \begin{cases}
  1-6\left(\frac{r}{h}\right)^2+6\left(\frac{r}{h}\right)^3,  &\quad 0\leq\frac{r}{h}\leq\frac{1}{2},\\
  2\left(1-\frac{r}{h}\right)^3,  &\quad \frac{1}{2} < \frac{r}{h} \leq 1,\\
  0,  &\quad \frac{r}{h} >1.\\
  \end{cases}
\end{equation}
The distribution of emissivities are then computed as a function of 
gas temperature and hydroden number density on each grid by bi-linearly 
interplating the grids of emissivities generated using CLOUDY 
({\bf version and citation}).
The temperature of the grids are in the range $?? < T < ??$K in intervals of
$\Delta \log_{10} T = ?.??$ and hydrogen number densities are in the range
$?? < n_H < ?? {\rm cm}^{-3}$ in intervals of $\Delta \log_{10} n_H = ??$.
The grids are computed at redshift z = 0 assuming gas is of solar metallicity,
optical thin and {\bf blabla}.

To compute the \hi{} neutral fraction, like \citep{Gutcke16}, we used the 
self-shielding approximation presented in \citep{Rahmati13}.
{\bf say more detail about it}. Contrastly, To compute the \ovi{}, the 
self-shielding approximation is not necessary to consider. {\bf reason}

\subsection{Luminosity Calculation}
\label{sec:lum}

In order to estimate the evolution of total \ovi{} and \hi{} luminosities inside
$2 \Rvir$, we employ the grids 
of emissivities computed since z=4 to present to evaluate the evolution of 
luminosities of all galaxies in NIHAO sample from particle data directly.
The method is similar to the one used by \citep{Sravan15}, the \ovi{} and \hi{} 
luminosities of each gas particle are calculated as
\begin{equation}
L = \epsilon_{\odot}\left(z,n_H,T\right)\left(\frac{m_{gas}}{\rho_{gas}}\right)
    \left(\frac{Z_{gas}}{Z_{\odot}}\right),
\end{equation}
where $m_{gas}$, $\rho_{gas}$ and $Z_{gas}$ are the mass, density and 
metallicity of the gas particle, and $\epsilon_{\odot}\left(z,n_H,T\right)$
is the emissivity interpolated bi-linearly from pre-computed grids of 
emissivities with given redshift, hydrogen number density and temperature.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 3 RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

\subsection{\protect\ovi{} and \protect\hi{} Emissivity Maps and Radial Profile}
\label{sec:map_rp}

\subsection{Evolution of Luminosities}
\label{sec:evo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 4: SUMMARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}
\label{sec:sum}

\section*{Acknowledgments}

{\sc Gasoline} was written by Tom Quinn and James Wadsley. Without
their contribution, this paper would have been impossible.
The simulations were performed on the {\sc theo} cluster of the
Max-Planck-Institut f\"ur Astronomie and the {\sc hydra} cluster at
the Rechenzentrum in Garching; and the Milky Way supercomputer, funded
by the Deutsche Forschungsgemeinschaft (DFG) through Collaborative
Research Center (SFB 881) "The Milky Way System" (subproject Z2),
hosted and co-funded by the J\"ulich Supercomputing Center (JSC). We
greatly appreciate the contributions of all these computing
allocations. AVM acknowledge support through the
Sonderforschungsbereich SFB 881 “The Milky Way System” (subproject A1)
of the German Research Foundation (DFG).  The analysis made use of the
pynbody package \citep{Pontzen13}.
%
The authors acknowledge support from the MPG-CAS through the
partnership programme between the MPIA group lead by AVM and the PMO
group lead by XK.
LW acknowledges support of the MPG-CAS student
programme.
XK acknowledge the support from NSFC project
No.11333008 and the ``Strategic Priority Research Program the Emergence
of Cosmological Structures'' of the CAS(No.XD09010000).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{}

%%AAAAAAAA

% Do Hot Haloes Around Galaxies Contain the Missing Baryons?
\bibitem[Anderson \& Bregman (2010)]{Anderson10} Anderson, M.~E., Bregman, J.~N.\ 2010, \apj, 714, 320

% Extended Hot Halos around Isolated Galaxies Observed in the ROSAT All-Sky Survey
\bibitem[Anderson et al.(2013)]{Anderson13} Anderson, M.~E., Bregman, J.~N., Dai, X.\ 2013, \apj, 762, 106

% Fundamental differences between SPH and grid methods
\bibitem[Agertz et al.(2007)]{Agertz07} Agertz, O., Moore, B., Stadel, J.\ 2007, \mnras, 380, 963

%%BBBBBBBB

% On the galaxy stellar mass function, the mass-metallicity relation and the implied baryonic mass function
\bibitem[Baldry et al.(2008)]{Baldry08} Baldry, I.~K., Glazebrook, K., Driver, S.~P.\ 2008, \mnras, 388, 945

%The Average Star Formation Histories of Galaxies in Dark Matter Halos
% from z = 0-8
\bibitem[Behroozi et al.(2013)]{Behroozi13} Behroozi, P.~S.,
  Wechsler, R.~H., \& Conroy, C.\ 2013, \apj, 770, 57

%A First Estimate of the Baryonic Mass Function of Galaxies
\bibitem[Bell et al.(2003)]{Bell03} Bell, E.~F., McIntosh, D.~H., Katz, N., Weinberg, M.~D.,\ 2003, \apj, 585, 117

% Multiscale Gaussian Random Fields and Their Application to Cosmological Simulations
%\bibitem[Bertschinger(2001)]{Bertschinger01} Bertschinger, E.\ 2001, 
%\apjs, 137, 1

% The Search for the Missing Baryons at Low Redshift
\bibitem[Bregman (2007)]{Bregman07}
Bregman, J.~N.\ 2007, ARAA, 45, 221

%% CCCCCCCC

%Where are the baryons
\bibitem[Cen \& Ostriker (1999)]{Cen09} 
Cen, R.~Y., Ostriker, J.~P.\ 1999, \apj, 514, 1

%% DDDDDDDD

%Dark halo response and the stellar initial mass function in early-type and late-type galaxies
\bibitem[Dutton et al.(2011)]{Dutton11} Dutton, A.~A., Conroy, 
  C., van den Bosch, F.~C., et al.\ 2011, \mnras, 416, 322

%The baryonic Tully-Fisher relation and galactic outflows
\bibitem[Dutton(2012)]{Dutton12} Dutton, A.~A.\ 2012,
  \mnras, 424, 3123

%Cold dark matter haloes in the Planck era: evolution of structural parameters for Einasto and NFW profiles
\bibitem[Dutton \& Macci{\`o}(2014)]{Dutton14} Dutton,
  A.~A., \& Macci{\`o}, A.~V.\ 2014, \mnras, 441, 3359

%% EEEEEEEEE

%% FFFFFFFFF

% On the Hot Gas Content of the Milky Way Halo
\bibitem[Fang et al.(2013)]{Fang13} Fang, T., Bullock, J., Boylan-Kolchin, M.\ 2013, \apj, 762, 20

%
\bibitem[Ford et al.(2013)]{Ford13} Ford, A.~B., Oppenheimer, B.~D., Dav{\`e}, R., et al.\ 2013, \mnras, 432, 89

%
\bibitem[Ford et al.(2015)]{Ford15} Ford, A.~B., Werk, J.~W., Dav{\`e}, R., et al.\ 2015, arXiv:1503.02084

% The Cosmic Baryon Budget
\bibitem[Fukugita et al.(1998)]{Fukugita98} Fukugita, M., Hogan, C.~J., Peebles, P.~J.~F.\ 1998, \apj, 503, 518

%% GGGGGGGGG

% A Huge Reservoir of Ionized Gas around the Milky Way: Accounting for the Missing Mass
\bibitem[Gupta et al. (2012)]{Gupta12} Gupta, A., Mathur, S., Krongold, Y., et al.\ 2012, \apj, 756L, 8

\bibitem[Gutcke et al.(2016)]{Gutcke16} Gutcke, T.~A., Stinson, G.~S., Macci{\`o}, A.~V., et al.\ 2016, arXiv:1602.06956

%% HHHHHHHHH

%
\bibitem[Hummels et al.(2013)]{Hummels13} Hummels, C.~B., Bryan, G.~L., Smith, B.~D., et al.\ 2013, \mnras, 430,1548

%% JJJJJJJJJJJ

%% KKKKKKKKKKK

%A superbubble feedback model for galaxy simulations
\bibitem[Keller et al.(2014)]{Keller14} Keller, B.~W., Wadsley, 
  J., Benincasa, S.~M., \& Couchman, H.~M.~P.\ 2014, \mnras, 442, 3013

% Stellar mass -- halo mass relation and star formation efficiency in
% high-mass halos
\bibitem[Kravtsov et al.(2014)]{Kravtsov14} Kravtsov, A., 
Vikhlinin, A., \& Meshscheryakov, A.\ 2014, arXiv:1401.7329

%% LLLLLLLLL

%% MMMMMMMMM

% The Baryon Content of Cosmic Structures
\bibitem[McGaugh et al.(2010)]{McGaugh10} McGaugh, S.~S., 
Schombert, J.~M., de Blok, W.~J.~G., Zagursky, M.~J.\ 2010, \mnras,
708, 14

% A refined particle method for astrophysical problems
\bibitem[Monaghan \& Lattanzio(1985)]{Monaghan85} Monaghan, J.~J., Lattanzio, J.~C., 1985, \aap, 149, 135

% Galactic star formation and accretion histories from matching galaxies to dark matter haloes
\bibitem[Moster et al.(2013)]{Moster13} Moster, B.~P., Naab, T., 
\& White, S.~D.~M.\ 2013, \mnras, 428, 3121

%% NNNNNNNNN

%% OOOOOOOO

%
\bibitem[Oppenheimer et al,(2016)]{Oppenheimer16}
Oppenheimer, B.~D., Crain, R.~A., Schaye, J., et al.\ 2016, arXiv:1603.05984

%% PPPPPPPPP

%Planck 2013 results. XVI. Cosmological parameters
\bibitem[the Planck Collaboration et  al.(2014)]{Planck14}
  Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2014,
  \aap, 571, AA16

% A Budget and Accounting of Metals at z ~ 0: Results from the COS-Halos Survey
\bibitem[Peeples et al.(2014)]{Peeples14} Peeples, M.~S., Werk, J.~K., Tumlinson, J., et al.\ 2014, \apj, 786, 54

% Dark MaGICC: the effect of dark energy on disc galaxy formation. Cosmology does matter
%\bibitem[Penzo et al.(2014)]{Penzo14} Penzo, C., Macci{\`o}, 
%  A.~V., Casarini, L., Stinson, G.~S., \& Wadsley, J.\ 2014, \mnras, 442, 176

% The baryon content of the universe
\bibitem[Persic \& Salucci(1992)]{Persic92} Persic, M., Salucci, P.\ 1992, \mnras, 258, 14

%pynbody: N-Body/SPH analysis for python
\bibitem[Pontzen et al.(2013)]{Pontzen13} Pontzen, A., Ro{\v s}kar, R., Stinson, G., \& Woods, R.\ 2013, Astrophysics Source Code Library, 1305.002

%% RRRRRRRRRRR

%Impact of local stellar radiation on the H I column density distribution
\bibitem[Rahmati et al.(2013)]{Rahmati13} Rahmati, A., Schaye, J., Pawlik, A.~H., Rai{\v c}evi{\'c}, M., 2013, MNRAS, 431,2261

%% SSSSSSSSSSSS

% The enrichment of the intergalactic medium with adiabatic feedback - I. Metal cooling and metal diffusion
\bibitem[Shen et al.(2010)]{Shen10} Shen, S., Wadsley, J., 
\& Stinson, G.\ 2010, \mnras, 407, 1581

% The Baryon Census in a Multiphase Intergalactic Medium: 30% of the Baryons May Still be Missing
\bibitem[Shull et al.(2012)]{Shull12}
Shull, J.~M., Smith, B.~D., Danforth, C.~W.\ 2012, \apj, 759, 23

%
\bibitem[Shull (2014)]{Shull14}
Shull, J.~M.\ 2014, \apj, 784, 142

%Strongly Time-Variable Ultra-Violet Metal Line Emission from the Circum-Galactic Medium of High-Redshift Galaxies
\bibitem[Sravan et al.(2015)]{Sravan15} Sravan, N., Faucher-Gigu{\`e}re, C.~-A., van de Voort, F., et al.\ 2015, arXiv:1510.06410

%
\bibitem[Stinson et al.(2012)]{Stinson12} Stinson, G.~S., Brook, C., Prochaska, J.~X., et al.\ 2012, \mnras, 425,1270

%Making Galaxies In a Cosmological Context: the need for early stellar feedback
\bibitem[Stinson et al.(2013)]{Stinson13} Stinson, G.~S., Brook, 
C., Macci{\`o}, A.~V., et al.\ 2013, \mnras, 428, 129

%
\bibitem[Suresh et al.(2015b)]{Suresh15b} Suresh, J., Rubin, K.~H.~R., Kannan, R., et al.\ 2015, arXiv:1511.00687

%% TTTTTTTTTTT

% Not Dead Yet: Cool Circumgalactic Gas in the Halos of Early-type Galaxies
\bibitem[Thom et al.(2012)]{Thom12}
Thom, C., Tumlinson, J., Werk, J.~K.\ 2012, \apjl, 758, L41

% The Large, Oxygen-Rich Halos of Star-Forming Galaxies Are a Major Reservoir
% of Galactic Metals
\bibitem[Tumlinson et al.(2011)]{Tumlinson11} Tumlinson, J., Thom, C., Werk, J., et al.\ 2011, Science, 334, 948

% The COS-Halos Survey: Rationale, Design and a Census of Circumgalactic Neutral Hydrogen
\bibitem[Tumlinson et al.(2013)]{Tumlinson13} Tumlinson, J., Thom, C., Werk, J., et al.\ 2013, \apj, 777, 59

%% VVVVVVVVVVVV

%% WWWWWWWWWWWWWW

%Gasoline: a flexible, parallel implementation of TreeSPH
\bibitem[Wadsley et al.(2004)]{Wadsley04} Wadsley, J.~W., Stadel, 
J., \& Quinn, T.\ 2004, \na, 9, 137

%On the treatment of entropy mixing in numerical cosmology
\bibitem[Wadsley et al.(2008)]{Wadsley08} Wadsley, J.~W., 
Veeravalli, G., \& Couchman, H.~M.~P.\ 2008, \mnras, 387, 427

  %NIHAO project I: Reproducing the inefficiency of galaxy formation across cosmic time with a large sample of cosmological hydrodynamical simulations
\bibitem[Wang et al.(2015)]{Wang15} Wang, L., Dutton, A.~A.,  Stinson, G.~S., et al.\ 2015, \mnras, 454, 83

% The COS-Halos Survey: Keck LRIS and Magellan MagE Optical Spectroscopy
\bibitem[Werk et al.(2012)]{Werk12} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2012, \apjs, 198, 3

% The COS-Halos Survey: An Empirical Description of Metal-line Absorption in the Low-redshift Circumgalactic Medium
\bibitem[Werk et al.(2013)]{Werk13} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2013, \apjs, 204, 17

%The COS-Halos Survey: Physical Conditions and Baryonic Mass in the Low-redshift Circumgalactic Medium
\bibitem[Werk et al.(2014)]{Werk14} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2014, \apj, 792, 8

%% YYYYYYYYYYYY

% Galaxy Group in the SDSS DR4. III. The Luminosity and Stellar Mass Functions
\bibitem[Yang et al.(2009)]{Yang09} Yang, X., Mo, H.~J., van den Bosch, F.~C.\ 2009, \apj, 695, 900

%% ZZZZZZZZZZZZZ

\end{thebibliography}

\label{lastpage}
\end{document}